Best practices for forecasting changes in biodiversity: an example using breeding birds

### 2017-08-18 manuscript/draft.Rmd

We evaluated 6 types of forecasting models (Table 1) at two time-scales. In the
first one, we divided our 32 years of data into 22 years of training data
and retained 10 years of data for evaluation via hindcasting. Second, we used
the full data set for training and made predictions through the year 2050. For
both time scales, we evaluated versions of each model with and without
correcting for observer effects, as described below.

@@

Observer effects are inherent in large data sets collected by different
observers, and are known to occur in BBS [@sauer1994]. For each forecasting
approach, we trained two versions of the corresponding model: one with
corrections for differences among observers, and one without (Figure 2). We
estimated the observer effects (and associated uncertainty about those effects)
with a linear mixed model, with observer as a random effect, built in the Stan
probabilistic programming language [@stan]. Because observer and site are
strongly related (observers tend to repeatedly sample the same site), site was
also included as a random effect to ensure that inferred deviations were
actually observer-related (as opposed to being related to the sites that a given
observer happened to see). The resulting model partitions the variance in
observed richness values into site-level variance, observer-level variance, and
residual variance (e.g.  variation within a site from year to year). The
site-level estimates can also be used directly as the "average" baseline model
(see below). The estimated observer effects can be subtracted from the richness
values for a particular observer to provide an estimate of how many species
would have been found by a "typical" observer. To incorporate uncertainty in
these "corrected" richness values into the forecasting models we collected 500
Monte Carlo samples from the model's posterior distribution, and fit each of the
downstream models with each of the Monte Carlo samples. Each Monte Carlo sample
represented a different possible set of observer-level and site-level random
effect values across the full 32-year dataset.

>>>>

We evaluated 6 types of forecasting models (Table 1) using hindcasting by
dividing the [30] years of data into [22] years of training data and [9] years
of data for evaluation. We also used the full data set to train all 6 models and
made annual forecasts through the year 2050. For both time scales, we made
forecasts using versions of each model with and without correcting for observer
effects, as described below.

@@

Observer effects are inherent in large data sets collected by different
observers, and are known to occur in BBS [@sauer1994]. For each forecasting
approach, we trained two versions of the corresponding model: one with
corrections for differences among observers, and one without (Figure 2). We
estimated the observer effects (and associated uncertainty about those effects)
with a linear mixed model, with observer as a random effect, built in the Stan
probabilistic programming language [@stan]. Because observer and site are
strongly related (observers tend to repeatedly sample the same site), site was
also included as a random effect to ensure that inferred deviations were
actually observer-related (as opposed to being related to the sites that a given
observer happened to see). The resulting model partitions the variance in
observed richness values into site-level variance, observer-level variance, and
residual variance (e.g.  variation within a site from year to year). The
site-level estimates can also be used directly as the "average" baseline model
(see below). The estimated observer effects can be subtracted from the richness
values for a particular observer to provide an estimate of how many species
would have been found by a "typical" observer. To incorporate uncertainty in
these "corrected" richness values into the forecasting models we collected 500
Monte Carlo samples from the model's posterior distribution, and fit each of the
downstream models with each of the Monte Carlo samples. Each Monte Carlo sample
represented a different possible set of observer-level and site-level random
effect values across the full 30-year dataset.
